Viscous Transport in Eroding Porous Media

### 2019-05-11 docs/transport/erosion_transport.tex
added a todo for Bryan
worked on numerics

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
Continuing with our previous work~\cite{qua-moo2018}, we use a completed
double-layer potential formulation for the velocity $\uu$.  The
double-layer potential is
\begin{align}
  \DDD[\eeta](\xx) = \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, $\nn$ is the unit outward
normal at $\yy$, and $\eeta$ is an unknown density function.  The Stokes
double-layer potential can be as the sum of the Laplace double-layer
potential and its gradient.  By identifying $\xx,\yy \in \RR^2$ with the
points $x,y \in \CC$, the Laplace double-layer potential for a
real-valued density function $\eta$ can also be written as the real part
of the Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \eta(\yy) ds_\yy = 
  \Re \left( \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy \right).
\end{align}
Therefore, the Stokes double-layer potential can be written in terms of
Cauchy integrals and its first derivative.  In addition, we also require
expressions for the derivative of the velocity field which can be
written in terms of a Cauchy integral and its first two derivatives.
Therefore, we introduce the notation
\begin{align}
  \label{eqn:cauchy}
  v[\eta](x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,  \quad
  v'[\eta](x) = -\frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{(x - y)^2} \, dy, \quad
  v''[\eta](x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{(x - y)^3} \, dy.
\end{align}
To simplify notation, whenever it will not cause confusion, we exclude
$\tau$ and instead write $v(x)$, $v'(x)$, and $v''(x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Barycentric Quadrature for $v''(x)$}
While accurate quadrature rules for $v(x)$ and $v'(x)$ have already been
developed~\cite{bar-wu-vee2015}, this work requires a similar quadrature
rule for $v''(x)$.  We mimic the approach developed
in~\cite{bar-wu-vee2015}.  First, if $x \in \Omega$ and $n \in
\mathbb{N}$, then 
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n > 1. \\
    \end{array}
  \right.
\end{align}
Applying this identity to the Cauchy integral theorem for $v''(x)$, we
have
\begin{align}
  0 &= \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x) \\
    &= \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - 
    \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{(y-x)^2 v''(x)}{(y-x)^3}\,dy -
    \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v(x)}{(y-x)^3}\, dy -
    \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2(y-x)v'(x)}{(y-x)^3}\, dy \\
    &= \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
Since we have constructed the integrand so that it is bounded for for
all $x \in \Omega$, the quadrature error of the trapezoid rule is
acceptable.  Applying the quadrature rule and solving for $v''(x)$, the
Barycentric quadrature rule for the second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}.
\end{align}
To compute $v(x)$ and $v'(x)$, the Barycentric quadrature rule in
equation~\eqref{eqn:vvpQuadrature} is used. If $x \in \Omega^c$, the
identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x} dy, 
\end{align}
is used, and the resulting Barycentric quadrature formula is
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Summation Methods}
\label{sec:fmm}
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
naively, it requires $\mathcal{O}(N^2)$ operations to directly.  Since
the shear stress must only be computed once per time step, and the
vorticity is only used as a post-processing step, we accept the
$\mathcal{O}(N^2)$ algorithm.  However, the velocity must be computed at
every GMRES iteration, so a fast algorithm is used.

To compute the velocity with $\mathcal{O}(N)$ operations, we start by
using the trapezoid rule accelerated with the fast multipole method
(FMM). Since the trapezoid rule is inaccurate when $x$ is close to
$\bd\Omega$, we replace the trapezoid rule with the Barycentric
quadrature rule for such points.

This
cost can be reduced to $\mathcal{O}(N)$ by applying the fast multipole
method (FMM).  Rather than developing an FMM for layer potentials that
involve high-order derivatives of the Green's functions, we

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Boundary Integral Equation Formulation}
\label{s:formulation}
Continuing with our previous work~\cite{qua-moo2018}, we use a completed
double-layer potential formulation for the velocity $\uu$.  The
double-layer potential is
\begin{align}
  \DDD[\eeta](\xx) = \frac{1}{\pi}\int_{\bd\Omega} 
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) \, ds_\yy, \quad \xx \in \Omega,
  \label{eqn:stokesDLP}
\end{align}
where $\rr = \xx - \yy$, $\rho = \|\rr\|$, $\nn$ is the unit outward
normal at $\yy$, and $\eeta$ is an unknown density function.  The Stokes
double-layer potential can be as the sum of the Laplace double-layer
potential and its gradient.  By identifying $\xx,\yy \in \RR^2$ with the
points $x,y \in \CC$, the Laplace double-layer potential for a
real-valued density function $\eta$ can also be written as the real part
of the Cauchy integral
\begin{align}
  \DD[\eta](\xx) = \frac{1}{2\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \eta(\yy) ds_\yy = 
  \Re \left( \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy \right).
\end{align}
Therefore, the Stokes double-layer potential can be written in terms of
Cauchy integrals and its first derivative.  In addition, we also require
expressions for the derivative of the velocity field which can be
written in terms of a Cauchy integral and its first two derivatives.
Therefore, we introduce the notation
\todo[inline]{BQ: Need to discuss boundary data and how to get it from
density $\eta$}
\begin{align}
  \label{eqn:cauchy}
  v[\eta](x) = \frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{x - y} \, dy,  \quad
  v'[\eta](x) = -\frac{1}{2\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{(x - y)^2} \, dy, \quad
  v''[\eta](x) = \frac{1}{\pi i} \int_{\bd\Omega}
    \frac{\eta(y)}{(x - y)^3} \, dy.
\end{align}
To simplify notation, whenever it will not cause confusion, we exclude
$\tau$ and instead write $v(x)$, $v'(x)$, and $v''(x)$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Barycentric Quadrature for $v''(x)$}
While accurate quadrature rules for $v(x)$ and $v'(x)$ have already been
developed~\cite{bar-wu-vee2015}, this work requires a similar quadrature
rule for $v''(x)$.  We mimic the approach developed
in~\cite{bar-wu-vee2015}.  First, if $x \in \Omega$ and $n \in
\mathbb{N}$, then 
\begin{align}
  \frac{1}{2\pi i}\int_{\bd\Omega} \frac{1}{(y-x)^n}\, dy = 
  \left\{
    \begin{array}{ll}
      1, & n = 1, \\
      0, & n > 1. \\
    \end{array}
  \right.
\end{align}
Applying this identity to the Cauchy integral theorem for $v''(x)$, we
have
\begin{align}
  0 &= \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y)}{(y-x)^3}\,dy - v''(x) \\
%    &= \frac{1}{2\pi i} \int_{\bd\Omega} 
%      \frac{2v^{-}(y)}{(y-x)^3}\,dy - 
%    \frac{1}{2\pi i} \int_{\bd\Omega} 
%      \frac{(y-x)^2 v''(x)}{(y-x)^3}\,dy -
%    \frac{1}{2\pi i} \int_{\bd\Omega} 
%      \frac{2v(x)}{(y-x)^3}\, dy -
%    \frac{1}{2\pi i} \int_{\bd\Omega} 
%      \frac{2(y-x)v'(x)}{(y-x)^3}\, dy \\
    &= \frac{1}{2\pi i} \int_{\bd\Omega} 
      \frac{2v^{-}(y) - v''(x)(y-x)^2 - 2v(x) - 2(y-x)v'(x)}
      {(y-x)^3}\, dy.
\end{align}
Since we have constructed the integrand so that it is bounded for for
all $x \in \Omega$, the quadrature error of the trapezoid rule is
acceptable.  Applying the quadrature rule and solving for $v''(x)$, the
Barycentric quadrature rule for the second derivative is
\begin{align}
  v''(x) &\approx \frac{2\sum\limits_{j=1}^N 
    \frac{v^{-}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}w_j}
    {\sum\limits_{j=1}^N \frac{1}{y_j-x}w_j}.
\end{align}
To compute $v(x)$ and $v'(x)$, the Barycentric quadrature rule in
equation~\eqref{eqn:vvpQuadrature} is used. If $x \in \Omega^c$, the
identity
\begin{align}
\frac{1}{x-a} &= -\frac{1}{2\pi i}\int_{\bd\Omega} 
    \frac{(y-a)^{-1}}{y-x} dy, 
\end{align}
is used, and the resulting Barycentric quadrature formula is
\begin{align}
  v''(x) &\approx \frac{1}{x-a}\frac{2\sum\limits_{j=1}^N
    \frac{v^{+}_{j} - v(x) - (y_j-x)v'(x)}{(y_j-x)^3}}
    {\sum\limits_{j=1}^N \frac{(y_j-a)^{-1}}{y_j-x}w_j}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Fast Summation Methods}
\label{sec:fmm}
The Barycentric quadrature rule achieves acceptable accuracy for
computing the velocity, shear stress, and vorticity.  However, applied
directly, it requires $\mathcal{O}(N^2)$ operations.  Since the shear
stress must only be computed once per time step, and the vorticity is
only used as a post-processing step, we accept the $\mathcal{O}(N^2)$
algorithm.  However, the velocity must be computed at every GMRES
iteration, which would make the overall method impractical without a
fast algorithm.

To compute the velocity with $\mathcal{O}(N)$ operations, we start by
using the trapezoid rule accelerated with the fast multipole method
(FMM) \todo[inline]{reference here}. For target points that are
sufficiently far from $\bd\Omega$, the resulting velocity is
sufficiently accurate.  However, the velocity at points that are too
close to $\bd\Omega$ must be corrected.  For a general point $\xx$,
\begin{align}
  \vv(\xx) = \int_{\Gamma} D(\xx,\yy) \eeta(\yy)\, ds_\yy + 
    \sum_{k=1}^{M} \int_{\gamma_k} D(\xx,\yy) \eeta(\yy)\, ds_\yy.
\end{align}
Assuming that $\xx$ is too close to $\gamma_i$, we first subtract the
velocity contribution of $\vv(\xx)$ due to $\gamma_i$.  Then, the
Barycentric quadrature rule is used to compute the velocity due to
$\gamma_i$ with more accuracy.  This is correction is applied to all
points that are close to an eroding body or the outer wall $\Gamma$.

The same strategy of using the trapezoid rule and then correcting with
the Barycentric quadrature is used to compute the shear stress on
$\bd\Omega$ and vorticity in $\Omega$.  The trapezoid rule is applied
with a direct summation, but this cost is only incurred once per time
step, and the majority of the computational effort is still required to
compute the velocity using the fast summation method.

To determine if a point $\xx \in \gamma_i$ needs to be corrected with
the Barycentric quadrature rule, we first estimate the proximity of
$\xx$ to $\bd\Omega$.  The criteria depends on the length of the outer
wall, $L_w$, the length of the eroding bodies $L_i$, $i=1,\ldots,M$, and
the number of discretization points $N_\out$ and $N_\iin$.
